Predictive Process Monitoring in Apromore

### 2018-04-09 CAiSE18-forum/apromore.tex
camera-ready version

Apromore is a Web-based advanced process analytics platform, developed by the BPM community under an open-source initiative.\footnote{\url{http://apromore.org}} Apromore was originally conceived as an advanced process model repository. However, today it offers a wide range of features which go beyond those for managing large process model collections, and include a variety of state-of-the-art process mining techniques. These are techniques for the automated discovery of BPMN models, for the conformance checking of BPMN models against event logs, the replaying of event logs on top of BPMN models, the detection and characterization of process drifts from event logs, the visual analysis of process performance, and many others.

All these features are exposed through a Web portal, and organized according to the phases of the BPM lifecycle: discovery, analysis, redesign, implementation and monitoring. These features can also be accessed as external Web services by third-party BPM software environments, such as ProM (for process mining) and WoPeD (for process modeling and verification).

>>>>

Apromore is a Web-based advanced process analytics platform, developed by the business process management (BPM) community under an open-source initiative. Apromore was originally conceived as an advanced process model repository. However, today it offers a wide range of features which go beyond those for managing large process model collections, and include a variety of state-of-the-art process mining techniques. These are techniques for the automated discovery of BPMN models, for the conformance checking of BPMN models against event logs, the replaying of event logs on top of BPMN models, the detection and characterization of process drifts from event logs, the visual analysis of process performance, and many others.

All these features are exposed through a Web portal, and organized according to the phases of the BPM lifecycle: discovery, analysis, redesign, implementation and monitoring \cite{Dumas2018}. These features can also be accessed as external Web services by third-party BPM software environments, such as ProM (for process mining) and WoPeD (for process modeling and verification).

### 2018-04-09 CAiSE18-forum/paper.tex
camera-ready version

%As an example, consider a doctor who needs to choose the most appropriate therapy for a patient. Historical data referring to patients with similar characteristics can be used to predict what therapy will be the most effective one and to advise the doctor accordingly. Meanwhile, in the context of a business process for managing loan applications, the applicant can be advised on the combinations of the loan amount and the length period for the loan that are the most likely to lead to acceptance of the application, given contextual information about the application and the personal data of the applicant (e.g., age, salary, etc.).
As an example, consider a business process for managing loan applications; the applicant can be advised about the combinations of loan amount and duration that are the most likely to lead to an acceptance of the application, given the history of the application until now and the personal data of the applicant (e.g., age, salary, etc.).

A range of approaches have been proposed in the literature to tackle common predictive process monitoring tasks \cite{Chamorro2017}.
However, these approaches have largely remained in the academic domain and have not been widely applied in real-time scenarios where users require a continuous predictive support.

To fill this gap, last year we started the \emph{Nirdizati} open-source project, aimed at developing a Web-based tool for the predictive monitoring of business processes. In this paper, we document the developments of Nirdizati since its creation, and its integration into the process analytics platform Apromore.

@@

Firstly, when a user uploads their log, the tool extracts and categorizes data attributes of the log. In order to properly construct feature vectors from business process traces, the attributes need to be categorized into static case attributes and dynamic event attributes. To this end, for each attribute in the log, we check whether it changes throughout the case lifetime. On the other hand, each attribute needs to be designated as either numeric or categorical.

@@

In order to that, we count the number of unique levels in the values of each attribute and if it less than a certain threshold, we decide the attribute to be categorical, otherwise it is numeric. These procedures are performed automatically upon the log uploading. Nevertheless, the user is given an option to override the automatic attribute definitions. %(Figure~\ref{fig:attribute-selector})
The resulting definitions are saved in a configuration file in a JSON format (Fig.~\ref{fig:listing}). Secondly, the log is internally split into training and validation set in a 80-20 proportion. The former is used to train the model, while the latter is used to evaluate the predictive power of the model. Next, all traces of a business process need to be represented as fixed-size feature vectors in order to train a predictive model. To this end, several encoding techniques were proposed in \cite{Leontjeva2015} and further refined in \cite{Teinemaa2017}, out of which we support four, namely last state encoding, frequency (aggregation) encoding, combined encoding and lossless index-based encoding. 
\begin{figure}[htp!]%[H]
	\centering
	 \includegraphics[width=.7\textwidth]{img/Listing}
	%\includegraphics[width=0.7\textwidth]{img/nirdizati-training}
	\caption{Example training configuration file.}
	\label{fig:listing}
	\vspace{-\baselineskip}
\end{figure}

@@

The status of the trained models can be verified using the collapsible drawer in the right-hand corner. Upon the training completion, a serialized Python object in the pickle format is produced. It describes a trained predictive model and which includes:

@@

The pipeline is built on top of the open-source Apache Kafka stream processing platform. The ``predictor'' components of the pipeline are the predictive models from the Training plugin bundle. The ``topic'' components are network-accessible queues of JSON messages with publisher/subscriber support. This allows the computationally intense work of the predictors to be distributed across a cluster of networked computers, providing scalability and fault-tolerance.
The ``collator'' component accumulates the sequence of events-to-date for each case, such that the prediction is a stateless function of the trained predictive model and of the case history.  This statelessness is what allows the predictors to be freely duplicated and distributed.  The ``joiner'' component composes the original events with the various predictions, ready for display on the dashboard.

@@

Video demos of the model training and of the runtime functionality can be found at \url{http://youtu.be/CYwLTKAbF1Q} and at \url{http://youtu.be/Q4WVebqJzUI}. The source code is available under the LGPL version 3.0 license at \url{https://github.com/apromore/ApromoreCode}.
%\vspace{-\baselineskip}
\bibliographystyle{splncs03}
\bibliography{paper}

>>>>

%As an example, consider a doctor who needs to choose the most appropriate therapy for a patient. Historical data referring to patients with similar characteristics can be used to predict what therapy will be the most effective one and to advise the doctor accordingly. Meanwhile, in the context of a business process for managing loan applications, the applicant can be advised on the combinations of the loan amount and the length period for the loan that are the most likely to lead to acceptance of the application, given contextual information about the application and the personal data of the applicant (e.g., age, salary, etc.).
There are many scenarios where it is useful to have reliable predictions. For example, in a purchase-to-pay business process, a customer can be advised about the estimated supply date. In case this date violates service-level agreement (SLA), preemptive actions can be taken by the supplier to prevent or mitigate the delay.

A range of approaches have been proposed in the literature to tackle common predictive process monitoring tasks. A recent survey of this field \cite{Chamorro2017} identified 39 distinct proposals (excluding subsumed ones) targeting time-related predictions, future path predictions and case outcome predictions.
However, these approaches have largely remained in the academic domain and have not been widely applied in real-time scenarios where users require a continuous predictive support.

To fill this gap, last year we started the \emph{Nirdizati} open-source project, aimed at developing a Web-based tool for the predictive monitoring of business processes \cite{Jorbina2017}. This paper documents a major revision of the project. Following state of the art in predictive processing monitoring and numerous feedback we received from prospective users, we have redesigned the user interface, technology stack and predictive techniques behind the tool. Furthermore, both training and runtime functionality have been integrated into the process analytics platform Apromore.\footnote{\url{http://apromore.org}}

@@

Firstly, when a user uploads their log, the tool extracts and categorizes data attributes of the log into static case attributes and dynamic event attributes. To this end, for each attribute in the log, we check whether it changes throughout the case lifetime. On the other hand, each attribute needs to be designated as either numeric or categorical.

@@

In order to do that, we count the number of unique levels in the values of each attribute and if it less than a certain threshold, we decide the attribute to be categorical, otherwise it is numeric. These procedures are performed automatically upon the log uploading. Nevertheless, the user is given an option to override the automatic attribute definitions. Proper attribute categorization ensures best training data quality. 
The resulting definitions are saved in a configuration file in a JSON format (Fig.~\ref{fig:listing}). Secondly, the log is internally split into training and validation set in a 80-20 proportion. The former is used to train the model, while the latter is used to evaluate the predictive power of the model. Next, all traces of a business process need to be represented as fixed-size feature vectors in order to train a predictive model. To this end, several encoding techniques were proposed in \cite{Leontjeva2015} and further refined in \cite{Teinemaa2017}, out of which we support four, namely last state encoding, frequency (aggregation) encoding, combined encoding and lossless index-based encoding. 
\begin{figure}[htp!]%[H]
	\centering
	 \includegraphics[width=.7\textwidth]{img/Listing}
	%\includegraphics[width=0.7\textwidth]{img/nirdizati-training}
	\caption{Example training configuration file.}
	\label{fig:listing}
	\vspace{-\baselineskip}
\end{figure}

@@

The status of the trained models can be verified using the collapsible drawer in the right-hand corner. Upon the training completion, a serialized Python object in the pickle format is produced. It describes a trained predictive model and includes:

@@

The pipeline is built on top of the open-source Apache Kafka stream processing platform.\footnote{\url{https://kafka.apache.org}} The ``predictor'' components of the pipeline are the predictive models from the Training plugin bundle. The ``topic'' components are network-accessible queues of JSON messages with publisher/subscriber support. This allows the computationally intense work of the predictors to be distributed across a cluster of networked computers, providing scalability and fault-tolerance.
The ``collator'' component accumulates the sequence of events-to-date for each case, such that the prediction is a stateless function of the trained predictive model and of the case history.  This statelessness is what allows the predictors to be freely duplicated and distributed.  The ``joiner'' component composes the original events with the various predictions, ready for display on the dashboard.

@@

Video demos of the model training and of the runtime functionality can be found at \url{http://youtu.be/CYwLTKAbF1Q} and at \url{http://youtu.be/Q4WVebqJzUI} respectively. The source code is available under the LGPL version 3.0 license at \url{https://github.com/apromore/ApromoreCode}.
%\vspace{-\baselineskip}
\bibliographystyle{splncs03}
\bibliography{paper}
